### 2013-01-06 edupar-13/410.tex
Talk about vanish/vanish troubles

\subsubsection{Pebbles Kernel}
For the fourth project, two-student teams
implement a kernel which complies to the
same specification as the reference kernel
they previously relied on.
They implement some approach to
synchronizing and blocking threads while
they are in kernel space,
a simple round-robin scheduler,
basic virtual memory,
a program loader,
code to handle various x86 exceptions,
and code for setting up and tearing down
threads and processes.
Students re-use the basic device drivers
they implemented in the second project.
%% TODO Space permitting, maybe have a table of system calls?
%% The 410 section is otherwise lacking in visual aids. -- bblum
These ingredients are combined to yield
twenty-five system calls
(life-cycle, thread management, memory management,
console I/O, and miscellaneous),
which are summarized in Figure~\ref{fig:syscalls}.
%% TODO Justify by mentioning average lines of code. -- bblum
For most students in the class, this is the
largest and most complicated software artifact they
have produced.
%% TODO Huh? I don't get how the sentence below is supposed to flow.
%% Seems like it wants to elaborate on the concerns, but the next sentence
%% is different. -- bblum
Because the test suite and the grading criteria
emphasize robustness and preemptibility of
kernel code,
there are many cross-cutting concerns.
Because the students are responsible for ensuring
the run-time invariants underlying all compiler-generated
code in the system (kernel and user-space),
they gain experience with debugging at both the
``algorithm level'' and the register/bit-field level.

>>>>

\subsubsection{Pebbles Kernel}
For the fourth project, two-student teams
implement a kernel which complies to the
same specification as the reference kernel
they previously relied on.
They implement some approach to
synchronizing and blocking threads while
they are in kernel space,
a simple round-robin scheduler,
basic virtual memory,
a program loader,
code to handle various x86 exceptions,
and code for setting up and tearing down
threads and processes.
Students re-use the basic device drivers
they implemented in the second project.
%% TODO Space permitting, maybe have a table of system calls?
%% The 410 section is otherwise lacking in visual aids. -- bblum
These ingredients are combined to yield
twenty-five system calls
(life-cycle, thread management, memory management,
console I/O, and miscellaneous),
which are summarized in Figure~\ref{fig:syscalls}.

%% TODO Justify by mentioning average lines of code. -- bblum
For most students in the class, this is the
largest and most complicated software artifact they
have produced.
%% TODO Huh? I don't get how the sentence below is supposed to flow.
%% Seems like it wants to elaborate on the concerns, but the next sentence
%% is different. -- bblum
Because the test suite and the grading criteria
emphasize robustness and preemptibility of
kernel code,
there are many cross-cutting concerns.
%
%% TODO: Dave, is it OK to say the following sentence? -- bblum
Widely regarded as the most difficult concurrency problem in the project
is that of coordinating a parent and a child task that ``simultaneously''
invoke \x{vanish()}: the parent task must reassign ownership of the child to
\x{init}, the root process, while the child task must check its parent's state
for any \x{wait()}ing threads which need to be woken up.
Many solutions, due to design constraints imposed by other parts of the kernel,
are prone to data races or deadlocks without careful thought.
%
Because the students are responsible for ensuring
the run-time invariants underlying all compiler-generated
code in the system (kernel and user-space),
they gain experience with debugging at both the
``algorithm level'' and the register/bit-field level.

### 2013-01-06 edupar-13/intro.tex
write a ton more
Talk about vanish/vanish troubles

In this paper, we describe our experience teaching concurrent programming at the operating systems level to advanced undergraduate students at Carnegie Mellon University (CMU).
In 15-410, the Operating Systems Design and Implementation class at CMU, students complete an aggressively challenging kernel project, in which they implement threads and processes in code that will boot on real hardware.
Presented with a kernel specification, called Pebbles~\cite{kspec}, students build a threading and synchronization library that runs in userspace on top of a Pebbles reference kernel in a two-week project, and subsequently, in a 6-week project, write their own kernel implementation.
Finally, in another two-week project, students implement a small extension to their Pebbles kernel. This project changes every semester, and in the past has included extending the kernel to be able to hypervise other kernels and to be able to run in SMP systems.

@@

The rest of this paper is organized as follows.
Section~\ref{sec:projects} gives an overview of 15-410's five programming projects.
Section~\ref{sec:pebbles} discusses the Pebbles specification in detail.
Section~\ref{sec:landslide} describes Landslide's testing mechanisms.
Section~\ref{sec:future} concludes, discussing strategies for making Landslide more accessible to struggling students, and how Landslide might be applied to related operating systems projects at other universities.

>>>>

In this paper, we describe our experience teaching concurrent programming at the operating systems level to advanced undergraduate students at Carnegie Mellon University (CMU).
In 15-410, the Operating Systems Design and Implementation class at CMU, students complete an aggressively challenging kernel project, in which they implement threads and processes in code that will boot on real hardware.
Presented with a kernel specification, called Pebbles~\cite{kspec}, students build a threading and synchronization library that runs in userspace on top of a Pebbles reference kernel in a two-week project, and subsequently, in a 6-week project, write their own kernel implementation.
During these projects students face very demanding synchronization problems, including how to implement locking primitives and how to coordinate a parent and a child process that are exiting simultaneously.
Finally, in another two-week project, students implement a small extension to their Pebbles kernel. This project changes every semester, and in the past has included extending the kernel to be able to hypervise other kernels and to be able to run in SMP systems.

@@

The rest of this paper is organized as follows.
Section~\ref{sec:curriculum} gives an overview of 15-410's five programming projects.
Section~\ref{sec:pebbles} discusses the Pebbles specification in detail.
Section~\ref{sec:landslide} describes Landslide's testing mechanisms.
Section~\ref{sec:eval} describes how we worked with students to evaluate Landslide.
Section~\ref{sec:future} concludes, discussing strategies for making Landslide more accessible to struggling students, and how Landslide might be applied to related operating systems projects at other universities.
